---
title: "General Social Survey"
author: "Karla Cota Lopez"
date: "Sept 10, 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Question: is gun ownership independent of opinion on capital punishment?

$H_0$: gun ownership and opinion on capital punishment are not related

$H_A$: gun ownership and opinion on capital punishment are related

OWNGUN ~ CAPPUN both are categorical, 
thus i use cat~cat

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")

```

This barchart shows gun owners are more likely to favor capital punishment

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~OWNGUN + CAPPUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

This numerical summary shows that those who own guns favor capital punishment

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

We reject the null hypothesis because the P-value is under 0.05 and conclude that they are related.
The Fisher exact test shows that statistically speaking if you own a gun you are 2.3 times more likely to favor capital punishment.

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Question: is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$:There is no difference in years of education between those that did vote and those that didn't in 2012.

$H_A$:There is a difference in years of education between those that did vote and those that didn't in 2012.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
boxplot(EDUC~VOTE12,data=GSSdata)
```

The Graph shows that people who voted in 2012 tended to have more education than those who didn't vote in 2012.

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```

The average for years of education for those who voted in 2012 is 15.7 years. People who did not vote had an average of 13.4 years of education.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```

We reject the null hypothesis, there is a difference in education in those who voted in 2012 and in those who didn't.

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Question: Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

$H_0$: A persons performance on the wordsum test is not independent of their level of schooling.

$H_A$:A persons performance on the wordsum test is independent of their level of schooling.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
boxplot(WORDSUM~DEGREE,data=GSSdata)
```

This boxplot shows us the different wordsum scores for each degree. Highscool has the lowest, junior college balances out between low scores and high scores and the graduate degree seems to have the highest wordsum scores.


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

This table shows us the average scores of each degree. The average for LTHS being 6.8, average for junior college is 8.9, and the graduate degree is 9.5. 
Therefore this concludes that those with graduate degrees get the highest wordsum scores.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSSdata)
summary.aov(model1)
```

We reject the null hypothesis. The p-value is 0.000236 which is under 0.05. There is a difference between the scores on wordsum depending on their schooling degree. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

To conclude, all questions had a relationship, we rejected the null hypothesis for all 3 questions. All the P-value results were under 0.05. Those who own guns are more likely to be in favor or capital punishment. If you have a higher education you are more likely to vote. Also if you have a higher education you are more likely to get a higher wordsum score. 
